\texttt{DreamCoder}: Growing libraries of concepts with wake-sleep program induction

### 2018-10-06 pp.tex
last-minute tweaks

\begin{SaveVerbatim}[]{ListSketch}
Bool ::= (<= Int) | (>= Int)
       | (= Int)
Int  ::= 0
       | (+1 Int) | (-1 Int)
       | (length List)
       | (head List)
List ::= nil | X
       | (filter Bool List)
       | (tail List)
       | (list Int)
\end{SaveVerbatim}

\begin{SaveVerbatim}[]{ListSketchFull}
Program ::= (if (Bool Int)
                List
                (append RecursiveList
                        RecursiveList
                        RecursiveList))
RecursiveList ::= List
                | (recurse List)
Bool ::= (<= Int) | (>= Int)
       | (= Int)
Int  ::= 0
       | (+1 Int) | (-1 Int)
       | (length List)
       | (head List)
List ::= nil | X
       | (filter Bool List)
       | (tail List)
       | (list Int)
\end{SaveVerbatim}

\begin{SaveVerbatim}[]{Reverse}
(if (<= (+1 0) (length X))
    X
    (append (recurse (tail X))
            (list (head X))))
\end{SaveVerbatim}
\begin{SaveVerbatim}[]{Reversep}
(if (= (+1 0) (length X))
    X
    (append (recurse (filter (<= (+1 0)) (tail X)))
            (list (head X))))
\end{SaveVerbatim}

\begin{SaveVerbatim}[]{Count}
(length (filter (= (head X)) (tail X)))
\end{SaveVerbatim}

\begin{SaveVerbatim}[]{Countp}
(-1 (length (filter (= (head X)) X)))
\end{SaveVerbatim}

\begin{SaveVerbatim}[]{Sort}
(if (<= (+1 0) (length X))
    X
    (append (recurse (filter (<= (head X)) (tail X)))
            (list (head X))
            (recurse (filter (>= (head X)) (tail X)))))
\end{SaveVerbatim}

\begin{SaveVerbatim}[]{Sortp}
(if (= (+1 0) (length X))
    X
    (append (recurse (tail X))
            (list (length X))))
\end{SaveVerbatim}

\begin{SaveVerbatim}[]{TextSketch}
Program ::= Term
          | Program + Term
Term    ::= String
          | substr(Pos,Pos)
Pos     ::= Number
          | pos(Str,Str,Num)
Num     ::= 0 | 1 | 2 | ... 
          | -1 | -2 | ...
Str     ::= Character 
          | Character + String
Character ::= a | b | c | ...
\end{SaveVerbatim}

\newcommand{\theSystem}{\textsc{ProgramSample}}
\usepackage{arydshln}

@@

\begin{frame}{Human program induction everywhere}
\begin{tikzpicture}
  \node(Lisp) at (0,0) {\includegraphics[width = 5cm]{figures/1975.png}};
  \visible<2->{
    \node(colorless) at (6,0) {\includegraphics[width = 5cm]{figures/colorless.png}};
    \node at ([yshift=-2cm]colorless.south) {\includegraphics[width = 4cm]{figures/child.jpg}};
  }
  \visible<3->{
    \node(aqueduct) at ([yshift=-1cm,xshift=0cm]Lisp.south) {\includegraphics[width = 5cm]{figures/aqueduct.jpg}};
    \node(arch) at ([xshift=1cm,yshift=0cm]aqueduct.south) {\includegraphics[width = 1.5cm]{figures/arch.jpg}};
  }
  \visible<4->{
    \node at ([yshift=-1cm,xshift=-2cm]aqueduct.south) {\includegraphics[width = 2cm]{figures/tapestry.jpg}};
    \node(wiringDiagram) at ([yshift=-1cm,xshift=1.6cm]aqueduct.south) {\includegraphics[width = 5cm]{figures/wiringDiagram.jpg}};
  }
  \visible<5->{
    \node at ([xshift=2cm,yshift=1cm]wiringDiagram.east) {\includegraphics[width = 5cm]{tree.jpg}};
    }
  \end{tikzpicture}

@@

  \begin{itemize}
  \item Library of concepts (declarative knowledge; generative model over programs)
    \item Inference strategy (procedural knowledge)
  \end{itemize}
  \only<2>{
  \begin{tikzpicture}
    \node(problem) at (0,0) {\includegraphics[width = 2cm]{figures/functions/48.png}};
    \node(synthesizer)[draw,align=center] at ([xshift=3cm]problem.east) {Learned \\program inducer};
    \draw[->] (problem.east) -- (synthesizer.west);
    \node(program)[draw, align=center] at ([xshift=3cm]synthesizer.east) {program:\\$f(x) = 0.7x^2+1.2$};
    \draw[->] (synthesizer.east) -- (program.west);
  \end{tikzpicture}}
    \renewcommand\code\texttt
    \only<3>{
\Wider[4em]{  \begin{tikzpicture}
    \node[draw,align=center](problem) at (0,0) {
            \begin{tabular}{l}
          \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
          \code{[1\, 2\, 3\, 4]}$\to$\code{[3\, 4]} \\
          \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} \\
          \end{tabular}
};
    \node(synthesizer)[draw,align=center] at ([xshift=2.5cm]problem.east) {Learned \\program inducer};
    \draw[->] (problem.east) -- (synthesizer.west);
    \node(program)[draw, align=center] at ([xshift=2.5cm]synthesizer.east) {program:\\\code{(filter ($\lambda$ (x)}\\\code{ (> x 2)))}};
    \draw[->] (synthesizer.east) -- (program.west);
  \end{tikzpicture}

  \vspace{0.5cm}Concepts: \code{filter}, etc\\Inference strategy: neurosymbolic search for programs
    }}

@@

@@

@@

\begin{picture}(50,70) \put(250,0){\hbox{\includegraphics[width = 3cm]{figures/Lucas}}} \end{picture} 
\end{frame}
\begin{frame}{Learning the basics of programming}
  \includegraphics[width = \textwidth]{figures/McCarthy.png}

@@

\end{frame}
\begin{frame}{List functions \& Text editing: Learning curves on hold out tasks}

@@

%% \begin{frame}{Symbolic regression from visual input}
%% \centering\includegraphics[width = 5cm]{symbolicRegression.png}
%% \end{frame}

@@

\begin{frame}{Turtle graphics --- Learning curves}
  \centering
  \includegraphics[width=9cm]{figures/dreams/montageTalkLC.eps}
  %% \begin{itemize}
  %%   \item $\frac\pi2$ and $\frac\pi4$ from $\pi$, $2$, $+$ and $/$
  %%   \item A line of length \texttt{n} followed with a right angle
  %%   \item Loops of length \texttt{n} that uses the number \texttt{n} inside.
  %%   \item Unit line then teleport back to origin
  %%   \item ...
  %% \end{itemize}
\end{frame}

@@

  \only<1>{\centering\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_4.png}}
  \only<2>{\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_1.png}}
  \only<3>{\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_7.png}}
  \only<4>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_12.png}}
  \only<5>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_14.png}}
  \only<6>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_5.png}}
  \only<7>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_13.png}}

@@

   \underline{More human-like machine intelligence}\\%Flexibly adapting to new problem domains:
   \begin{itemize}
   \item    Acquiring a domain-specific representation (DSL)
     \item Learning  to write programs (recognition model)
   \end{itemize}
   DreamCoder: an algorithm for jointly realizing these goals

>>>>

\usepackage{arydshln}

@@

\begin{frame}{Human program induction everywhere}
\begin{tikzpicture}
  \node(Lisp) at (0,0) {\includegraphics[width = 5cm]{figures/1975.png}};
  \visible<2->{
    \node(colorless) at (6,0) {\includegraphics[width = 5cm]{figures/colorless.png}};
    \node at ([yshift=-2cm]colorless.south) {\includegraphics[width = 4cm]{figures/child.jpg}};
  }
  \visible<3->{
    \node(aqueduct) at ([yshift=-1cm,xshift=0cm]Lisp.south) {\includegraphics[width = 5cm]{figures/aqueduct.jpg}};
    \node(arch) at ([xshift=1cm,yshift=0cm]aqueduct.south) {\includegraphics[width = 1.5cm]{figures/arch.jpg}};
  }
  \visible<4->{
    \node at ([yshift=-1cm,xshift=-2cm]aqueduct.south) {\includegraphics[width = 2cm]{figures/tapestry.jpg}};
    \node(wiringDiagram) at ([yshift=-1cm,xshift=1.6cm]aqueduct.south) {\includegraphics[width = 5cm]{figures/wiringDiagram.jpg}};
  }
  %% \visible<5->{
  %%   \node at ([xshift=2cm,yshift=1cm]wiringDiagram.east) {\includegraphics[width = 5cm]{tree.jpg}};
  %%   }
  \end{tikzpicture}

@@

  \begin{itemize}
  \item Library of concepts (declarative knowledge; generative model over programs)
    \item Inference strategy (procedural knowledge)
  \end{itemize}
  \only<2>{
  \begin{tikzpicture}
    \node(problem) at (0,0) {\includegraphics[width = 2cm]{figures/functions/48.png}};
    \node(synthesizer)[draw,align=center] at ([xshift=3cm]problem.east) {Learned \\program inducer};
    \draw[->] (problem.east) -- (synthesizer.west);
    \node(program)[draw, align=center] at ([xshift=3cm]synthesizer.east) {program:\\$f(x) = 0.7x^2+1.2$};
    \draw[->] (synthesizer.east) -- (program.west);
  \end{tikzpicture}

    \vspace{0.2cm}Concepts: $x^2$, etc\\Inference strategy: neurosymbolic search for programs}
  \renewcommand\code\texttt
    \only<3>{
  \begin{tikzpicture}
    \node(problem) at (0,0) {\includegraphics[width = 2cm]{figures/logoTasks/logo43.png}};
    \node(synthesizer)[draw,align=center] at ([xshift=3cm]problem.east) {Learned \\program inducer};
    \draw[->] (problem.east) -- (synthesizer.west);
    \node(program)[draw, align=center] at ([xshift=3cm]synthesizer.east) {program:\\\code{(for (i 4)}\\ \code{ (circle i))}};
    \draw[->] (synthesizer.east) -- (program.west);
  \end{tikzpicture}

  \vspace{0.2cm}Concepts: \code{circle}, etc\\Inference strategy: neurosymbolic search for programs
    }
%%     \only<4>{
%% \Wider[4em]{  \begin{tikzpicture}
%%     \node[draw,align=center](problem) at (0,0) {
%%             \begin{tabular}{l}
%%           \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
%%           \code{[1\, 2\, 3\, 4]}$\to$\code{[3\, 4]} \\
%%           \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} \\
%%           \end{tabular}
%% };
%%     \node(synthesizer)[draw,align=center] at ([xshift=2.5cm]problem.east) {Learned \\program inducer};
%%     \draw[->] (problem.east) -- (synthesizer.west);
%%     \node(program)[draw, align=center] at ([xshift=2.5cm]synthesizer.east) {program:\\\code{(filter ($\lambda$ (x)}\\\code{ (> x 2)))}};
%%     \draw[->] (synthesizer.east) -- (program.west);
%%   \end{tikzpicture}

%%   \vspace{0.5cm}Concepts: \code{filter}, etc\\Inference strategy: neurosymbolic search for programs
%%     }}

@@

\vspace{0.5cm}

@@

\textbf{Dechter et al.}: Exploration-Compression. Inspiration for DreamCoder.

@@

\begin{picture}(50,70) \put(250,0){\hbox{\includegraphics[width = 3cm]{figures/Lucas}}} \end{picture} 
\end{frame}

@@

\begin{frame}{List functions \& Text editing: Learning curves on hold out tasks}

@@

\begin{frame}{Learning the fundamentals of programming}
  \includegraphics[width = \textwidth]{figures/McCarthy.png}

\begin{frame}{Symbolic regression from visual input}
\centering\includegraphics[width = 5cm]{symbolicRegression.png}
\end{frame}

@@

%% \begin{frame}{Turtle graphics --- Learning curves}
%%   \centering
%%   \includegraphics[width=9cm]{figures/dreams/montageTalkLC.eps}
%%   %% \begin{itemize}
%%   %%   \item $\frac\pi2$ and $\frac\pi4$ from $\pi$, $2$, $+$ and $/$
%%   %%   \item A line of length \texttt{n} followed with a right angle
%%   %%   \item Loops of length \texttt{n} that uses the number \texttt{n} inside.
%%   %%   \item Unit line then teleport back to origin
%%   %%   \item ...
%%   %% \end{itemize}
%% \end{frame}

@@

  \only<1>{\centering\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_4.png}}
  \only<2>{\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_1.png}}
  \only<3>{\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_7.png}}
  \only<4>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_12.png}}
  \only<5>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_14.png}}
  \only<6>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_5.png}}
%  \only<7>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_13.png}}

@@

   \underline{More human-like machine intelligence}\\%Flexibly adapting to new problem domains:
   \begin{itemize}
   \item    Acquiring a domain-specific representation (DSL)
     \item Learning  to use that representation (recognition model)
   \end{itemize}
   DreamCoder: an algorithm for jointly realizing these goals
